# 2025-11-13

## Dependencies and setup

```{r}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(khroma)
library(rstatix)
library(knitr)
library(kableExtra)
library(viridis)
# library(webshot2)
```

```{r}
# Build relative path from R's working directory to what is actually the working
# directory for this experiment.
experiment_slug <- "2025-11-13-nutrient-maxlife-fixedtime"
working_directory <- paste(
  "experiments",
  experiment_slug,
  "analysis",
  sep = "/"
)
```

```{r}
# Use the cowplot theme as our default for ggplots
theme_set(theme_cowplot())
# Create directory to save plots
plot_dir <- paste(
  working_directory,
  "plots",
  sep = "/"
)
dir.create(
  plot_dir,
  showWarnings = FALSE
)
```

### Load and preprocess data

```{r}
tasks <- c("NOT", "NAND", "OR_NOT", "AND", "OR", "AND_NOT", "NOR", "XOR", "EQU")
max_pop_size <- 100 * 100
threshold <- 0.025 * max_pop_size
```


Load time series data.

```{r}
time_data_path <- paste(
  working_directory,
  "data",
  "time_series.csv",
  sep = "/"
)
time_data <- read_csv(time_data_path)

# Do any universal preprocessing
time_data <- time_data %>%
  mutate(
    SEED = as.factor(SEED),
    HOST_MIN_CYCLES_BEFORE_REPRO = as.factor(HOST_MIN_CYCLES_BEFORE_REPRO),
    HOST_REPRO_RES = as.factor(HOST_REPRO_RES)
  ) %>%
  mutate(
    treatment = paste0(
      "MC=", HOST_MIN_CYCLES_BEFORE_REPRO,
      "_RES=", HOST_REPRO_RES
    )
  ) %>%
  mutate(
    treatment = as.factor(
      treatment
    )
  )
```


## Which runs had hosts and/or symbiont populations that went extinct?

```{r}
sym_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_hosted_sym_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_syms <- sym_min_pop_sizes %>%
  group_by(
    treatment
  ) %>%
  summarize(
    reps_with_extinct_syms = sum(OrgCounts_hosted_sym_count == 0)
  )

host_min_pop_sizes <- time_data %>%
  group_by(SEED) %>%
  slice_min(
    OrgCounts_host_count,
    n = 1
  ) %>%
  slice_min(
    update,
    n = 1
  )

reps_with_extinct_hosts <- host_min_pop_sizes %>%
  group_by(
    treatment
  ) %>%
  summarize(
    reps_with_extinct_hosts = sum(OrgCounts_host_count == 0)
  )

extinctions <- full_join(
  reps_with_extinct_syms,
  reps_with_extinct_hosts
)
extinctions <- extinctions %>%
  arrange(
    desc(reps_with_extinct_hosts), desc(reps_with_extinct_syms)
  )
extinction_table <- kable(extinctions) %>%
  kable_styling(latex_options = "striped")
save_kable(
  extinction_table,
  paste(
    plot_dir,
    "extinctions.pdf",
    sep = "/"
  )
)
```

## Task performance all time

### Hosts

Quantify whether tasks have been performed over course of entire run.

```{r}
# Pivot task data long:
# 1) Select only relevant columns
host_over_time_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    HOST_MIN_CYCLES_BEFORE_REPRO,
    HOST_REPRO_RES,
    update,
    Tasks_host_task_AND_NOT,
    Tasks_host_task_AND,
    Tasks_host_task_EQU,
    Tasks_host_task_NAND,
    Tasks_host_task_NOR,
    Tasks_host_task_NOT,
    Tasks_host_task_OR_NOT,
    Tasks_host_task_OR,
    Tasks_host_task_XOR
  ) %>%
  pivot_longer(
    c(
      Tasks_host_task_AND_NOT,
      Tasks_host_task_AND,
      Tasks_host_task_EQU,
      Tasks_host_task_NAND,
      Tasks_host_task_NOR,
      Tasks_host_task_NOT,
      Tasks_host_task_OR_NOT,
      Tasks_host_task_OR,
      Tasks_host_task_XOR
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "Tasks_host_task_", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  )

host_parent_tasks_all_time_by_task <- host_over_time_parent_task_data %>%
  # filter(update <= 100000) %>%
  group_by(
    SEED,
    treatment,
    HOST_MIN_CYCLES_BEFORE_REPRO,
    HOST_REPRO_RES,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    HOST_MIN_CYCLES_BEFORE_REPRO,
    HOST_REPRO_RES,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )

# parent_tasks_all_time_total <- parent_tasks_all_time_by_task %>%

```

```{r}

host_parent_tasks_all_time_by_task_plt <- host_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = task,
      fill = task,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 31)
  ) +
  facet_grid(
    HOST_MIN_CYCLES_BEFORE_REPRO ~ HOST_REPRO_RES,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_parent_tasks_all_time_by_task_plt.pdf", sep = "/"),
  plot = host_parent_tasks_all_time_by_task_plt,
  width = 15,
  height = 15
)
```

### Symbionts

Quantify whether tasks have been performed over course of entire run.

```{r}
# Pivot task data long:
# 1) Select only relevant columns
over_time_sym_parent_task_data <- time_data %>%
  select(
    SEED,
    treatment,
    HOST_MIN_CYCLES_BEFORE_REPRO,
    HOST_REPRO_RES,
    update,
    Tasks_sym_task_AND_NOT,
    Tasks_sym_task_AND,
    Tasks_sym_task_EQU,
    Tasks_sym_task_NAND,
    Tasks_sym_task_NOR,
    Tasks_sym_task_NOT,
    Tasks_sym_task_OR_NOT,
    Tasks_sym_task_OR,
    Tasks_sym_task_XOR
  ) %>%
  pivot_longer(
    c(
      Tasks_sym_task_AND_NOT,
      Tasks_sym_task_AND,
      Tasks_sym_task_EQU,
      Tasks_sym_task_NAND,
      Tasks_sym_task_NOR,
      Tasks_sym_task_NOT,
      Tasks_sym_task_OR_NOT,
      Tasks_sym_task_OR,
      Tasks_sym_task_XOR
    ),
    names_to = "task",
    values_to = "task_in_parent_count"
  ) %>%
  mutate(
    task = str_replace(task, "Tasks_sym_task_", "")
  ) %>%
  mutate(
    task_completed_by_parent = task_in_parent_count >= threshold,
    task = factor(
      task,
      levels = tasks,
      labels = tasks
    )
  )


sym_parent_tasks_all_time_by_task <- over_time_sym_parent_task_data %>%
  group_by(
    SEED,
    treatment,
    HOST_MIN_CYCLES_BEFORE_REPRO,
    HOST_REPRO_RES,
    task
  ) %>%
  summarize(
    n = n(),
    parent_task_completed_ever = any(task_completed_by_parent)
  ) %>%
  ungroup() %>%
  group_by(
    treatment,
    HOST_MIN_CYCLES_BEFORE_REPRO,
    HOST_REPRO_RES,
    task
  ) %>%
  summarize(
    n = n(),
    parent_tasks_completed_ever = sum(parent_task_completed_ever == TRUE)
  )

# parent_tasks_all_time_total <- parent_tasks_all_time_by_task %>%

```

```{r}

sym_parent_tasks_all_time_by_task_plt <- sym_parent_tasks_all_time_by_task %>%
  ggplot(
    aes(
      x = task,
      color = task,
      fill = task,
      y = parent_tasks_completed_ever
    )
  ) +
  geom_col(
    position = "dodge",
    alpha = 0.5
  ) +
  geom_text(
    aes(label = parent_tasks_completed_ever),
    position = position_dodge(width = 1),
    vjust = -0.5,
    size = 2
  ) +
  scale_x_discrete() +
  scale_y_continuous(
    limits = c(0, 31)
  ) +
  facet_grid(
    HOST_MIN_CYCLES_BEFORE_REPRO ~ HOST_REPRO_RES,
    axes = "all_x",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "sym_parent_tasks_all_time_by_task_plt.pdf", sep = "/"),
  plot = sym_parent_tasks_all_time_by_task_plt,
  width = 15,
  height = 15
)
```


## Population size

Average / median population size

```{r}
pop_size_summary <- time_data %>%
  group_by(
    SEED,
    treatment
  ) %>%
  summarize(
    host_pop_size_avg = mean(OrgCounts_host_count),
    host_pop_size_median = median(OrgCounts_host_count),
    host_pop_size_variance = var(OrgCounts_host_count),
    sym_pop_size_avg = mean(OrgCounts_hosted_sym_count),
    sym_pop_size_median = median(OrgCounts_hosted_sym_count),
    sym_pop_size_variance = var(OrgCounts_hosted_sym_count)
  )
```

```{r}
host_pop_size_avg_plt <- pop_size_summary %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = host_pop_size_avg)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_wrap(STRESS_FREQUENCY~stress_death_chance, labeller = label_both) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_pop_size_avg_plt.pdf", sep = "/"),
  plot = host_pop_size_avg_plt,
  width = 10,
  height = 10
)
```

```{r}
host_pop_size_variance_plt <- pop_size_summary %>%
  ggplot(
    aes(x = treatment, fill = treatment, y = host_pop_size_variance)
  ) +
  geom_point(
    mapping=aes(color = treatment),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete() +
  # facet_wrap(STRESS_FREQUENCY~stress_death_chance, labeller = label_both) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      angle = 60,
      hjust = 1
    )
  )
ggsave(
  file = paste(plot_dir, "host_pop_size_variance_plt.pdf", sep = "/"),
  plot = host_pop_size_variance_plt,
  width = 10,
  height = 10
)
```

## Phenotypic diversity

### Final update

```{r}
summary_data <- summary_data %>%
  mutate(
    only_credit_first = case_when(
      task_credit == "H0S0" ~ "Neither",
      task_credit == "H0S1" ~ "Symbionts",
      task_credit == "H1S0" ~ "Hosts",
      task_credit == "H1S1" ~ "Both"
    )
  ) %>%
  mutate(
    only_credit_first = as.factor(only_credit_first)
  ) %>%
  mutate(
    sym_condition = factor(
      sym_condition,
      levels = c(
        "no-syms",
        "parasite",
        "mutualist"
      )
    )
  )
host_parent_phenotypic_entropy_plt <- summary_data %>%
  ggplot(
    aes(
      x = sym_condition,
      color = sym_condition,
      fill = sym_condition,
      y = CurUpdate_host_parent_entropy_task_sets
    )
  ) +
  geom_point(
    mapping=aes(color = sym_condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_x_discrete() +
  facet_grid(
    interaction_multiplier ~ only_credit_first,
    axes = "all_y",
    scales = "free",
    labeller = label_both
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    ),
    panel.border = element_rect(colour = "black")
  )
ggsave(
  file = paste(plot_dir, "host_parent_phenotypic_entropy_plt.pdf", sep = "/"),
  plot = host_parent_phenotypic_entropy_plt,
  width = 15,
  height = 15
)
```